Preregistration for modelling in ecolgy and conservation

### 2020-06-01 index.Rmd
#20 update manuscript - first draft workshop analysis methods and add new todos to in situ eval and interface / shiny app.

@@

**Activity 1 - Capturing Workflows and Decision Steps**

@@

All workshop outputs can be found here (\<Link to Workshop Outputs OSF repository\>).

After the workshop, data sheets from activity 1 were collated and coded using atlas.TI.

The point of this activity is to build on the suggested workflows for both fieldwork and modelling to provide structure and content to the preregistration templates.

At the end of this exercise we expect to:

Have a complete draft workflow for both fieldwork and modelling and data-analysis that will form the basis of the final preregistration templates.

Part I: Identify Scientific Workflows for ecological modelling, both real/in-practice and idealised.

Objective: Identify common workflow.

Step 1: As a group, reflect on personal experience to identify unique decision-steps. Can a generalised / common workflow across individuals be derived?

Step 2: Search literature to to identify an idealised scientific workflow for ecological modelling research.

Step 3: Code workflows identified in step 1 using codes identified in step 3. Revise common workflow.

Part II: Translating Workflows into Preregistration Templates

We translated the workflows into the preregistration template by using the following ontology:

1\. Decision Phase - each has a "title" and "description" or "definition", steps are grouped under phases.

2\. Decision Step - each has a "title" and "description" or "definition" choices are grouped into 'steps'. Corresponds to an activity or set of activities in a scientific workflow.

3\. Choices ---\> each 'choice' corresponds to a uniquely numbered item on the PRT requiring a response from user. The style should be in the form of a directive or a question. E.g. "explain how you will do xyzzy". Or "What performance measure will you use for assessing 'goodness of fit'?" In addition, we turned to existing preregistration templates to guide the wording  of some items.

@@

@@

@@

!\[Living Preregistration Workflow - Research Workflow for implementing a 'living' preregistration in a modelling context](./figures/LivingPreregistration.png)

>>>>

**Identifying**

@@

**Activity 1 - Capturing Scientific Workflows and Decision Points**

@@

**1\. Synthesise Workshop Outputs: Identify Scientific Workflows for ecological modelling, both real/in-practice and idealised.**

The goal of this analysis task was to generate a complete draft workflow for both fieldwork and modelling and data-analysis that will form the basis of the final preregistration templates. We used a combination of inductive and deductive coding \<is this what it is called\> to analyse the workshop outputs and generate a model scientific workflow for model-based research that would constitute the presumed scientific workflow that would be used to guide the structure of the preregistration template. We collated and coded the workshop outputs using \<atlas.TI, insert software citation and version\>, \<a qualitative coding software\>. All workshop outputs can be found here (\<Link to Workshop Outputs OSF repository\>).

A preliminary idealised scientific workflow for ecological modelling research was synthesised after reviewing the literature on 'good modelling practice' and structured decision-making literature in ecological management. Steps in the workflow were coded and broken down into 'phases' consisting of 'steps' and 'substep' \<insert link to appendices or just present the final?\>. Each decision-point in the personal modelling workflows generated from activity one of the collaborative workshop were coded as belonging to one of these phases, steps and substeps. We revised the coding structure \<inductively (is this the word)\>, according to patterns and themes identified across multiple personal workflows. For example, we added the final phase \<name for last phase\>, which aimed to capture the fact that analysts usually present the results of their modelling to clients, decision-makers or other stakeholders that must interpret the evidence and are responsible for making management decisions. In this way, we identified a common or generalised workflow from different types of applied ecological modelling projects that synthesised both \<idealised norms and norms of practice, Pu et al., see other philosophy of sci refs\>. This workflow is used to provide structure and content to the preregistration template.

**2\. Translating Workflows into Preregistration Templates**

We translated the model-based workflow identified in part I of workshop analysis into a text-based preregistration template. Reflecting on existing templates \<link to OSF repository of templates \>, we identified common elements / units of preregistration templates that we mapped onto the generalised modelling workflow identified in part I of workshop output analysis. We synthesised the following ontologies:

1\. Decision Phase - each has a "title" and "description" or "definition", steps are grouped under phases. \<Note, that these are being scrapped from the template and are turned into tool-tips\>

2\. Decision Step - each has a "title" and "description" or "definition" choices are grouped into 'steps'. Corresponds to an activity or set of activities in a scientific workflow. \<again, these are being scrapped and turned into tool-tips.\>

3\. Choices ---\> each 'choice' corresponds to a uniquely numbered item on the PRT requiring a response from user. The style should be in the form of a directive or a question. E.g. "explain how you will do xyzzy". Or "What performance measure will you use for assessing 'goodness of fit'?"

**3\. Designing a new methodology for implementing preregistration in model-based research**

We analysed the outputs of Activity 2 in the collaborative workshop identifying barriers and solutions to undertaking preregistration in model-based research and to broader adoption of preregistration in the applied ecology and modeller communities of practice to design a preregistration procedure appropriate for this research domain. We considered such questions as:

\- What should a research workflow look like for a modeller in an applied ecology setting who wishes to preregister their work?

\- How do we accommodate the iterative cycle of model development into the preregistration process?

\- At what stage in a modelling or structured decision-making process should preregistration begin?

We then used 'in situ evaluation' to test and evaluate both the template and procedure design.

@@

\<WHAT DID WE EVALUATE\>

NOT JUST THE TEMPLATES THEMSELVES BUT ALSO THE PROCEDURE, and the INTERFACE. I.e. bc *in situ evaluation is used to inform the design of a new* interface *- the shiny app...*

@@

Reflecting on the in situ evaluation - shiny app

*And the in situ evaluation is used to inform the design of a new* interface *- the shiny app.*

@@

![Living Preregistration Workflow - Research Workflow for implementing a 'living' preregistration in a modelling context](./figures/LivingPreregistration.png)
